\documentclass{article}

\usepackage{amsmath}

\begin{document}

\section*{Problem 2-1}

\noindent\begin{enumerate}
	\item[\textbf{\textit{a.}}] Each insertion sort on a sublist of length $k$ will take $\Theta(k^2)$ time. There are $\frac{n}{k}$ sublists. Multiplying the amount of time it takes to sort each sublist by the number of sublists gives:
	\begin{eqnarray*}
		k^2 \frac{n}{k} \\
		nk
	\end{eqnarray*}
	Thus it takes time $\Theta(nk)$ to sort all of the sublists.

	\item[\textbf{\textit{b.}}] Assume that $n$ and $k$ are both powers of 2 and that $n < k$. We know that merging an $n$-element sublist takes time $\Theta(n)$ and that for any iteration $0 \leq i < \log_2(\frac{n}{k})$ there are $\frac{n}{2^i k}$ sublists. The amount of time it takes to merge the sublists in an iteration $i$ is the amount of time it takes to merge a sublist of length $2^{i+1} k$ multiplied by the number of sublists to merge:
	\begin{eqnarray*}
		2^{i+1} k \cdot \frac{n}{2^i k} \\
		2n \\
		\Theta(n)
	\end{eqnarray*}
	Since it takes time $\Theta(n)$ to merge all the sublists for an iteration $i$ and there are $\log_2(\frac{n}{k})$ iterations, multiplying the amount of time it takes to merge all of the sublists for an iteration $i$ by the number of iterations to perform gives:
	\begin{eqnarray*}
		n \cdot log_2(\frac{n}{k}) \\
		\Theta(n \log_2(\frac{n}{k}))
	\end{eqnarray*}

	\item[\textbf{\textit{c.}}] Suppose we were to set $k$ such that $k = n$ (thus running the Insertion Sort on the entire array). Then the worst-case run time would be:
	\begin{eqnarray*}
		\Theta(nk + n \log_2(\frac{n}{k})) \\
		\Theta(n^2 + n \log_2(\frac{n}{n})) \\
		\Theta(n^2 + n \log_2(1)) \\
		\Theta(n^2)
	\end{eqnarray*}
	\ldots which is the worst-case run time of Insertion Sort. Successively dividing this maximum value of $k$ by 2 for each iteration $i$ such that $k = \frac{n}{2^i}$ yields the analysis:
	\begin{eqnarray*}
		\Theta \left ( n \left ( \frac{n}{2^i} \right ) + n \log_2 \left ( \frac{n}{\frac{n}{2^i}} \right ) \right )
	\end{eqnarray*}
	For low values of $i$ this yields the worst-case run time of Insertion Sort, but using values of $i$ such that $k \rightarrow 1$ yields the analysis:
	\begin{eqnarray*}
		\Theta \left ( nk + n \log_2 \left ( \frac{n}{k} \right ) \right ) \\
		\Theta \left ( n(1) + n \log_2 \left ( \frac{n}{1} \right ) \right ) \\
		\Theta \left ( n + n \log_2 \left ( n \right ) \right ) \\
		\Theta \left (n \log_2 \left ( n \right ) \right )
	\end{eqnarray*}
	\ldots which is simply the worst-case runtime of merge sort. Thus the largest value of $k$ as a function of $n$ such that the modified algorithm has the same running time as standard merge sort is $k = 1$.

	\item[\textbf{\textit{d.}}] In practice, we should chose a value of $k$ such that the actual runtime of Insertion Sort on arrays of length $k$ and smaller is less than the runtime of Merge Sort.
\end{enumerate}

\end{document}
